Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2025-08-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
new formatting
air formatting

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",

@@

    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing", "Please reorder testing and re",

@@

    PatientID, Name, DOB, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right

@@

njh <- read.csv("./Data_Cleaned/njh_full_list.csv",

@@

    "", "Test not performed", "Insufficient quant.",

@@

    Patient.ID, Patient.Name, Birth.Date, Genotype.1, Genotype.2,
    Sex, Age.at.test..in.years., Sweat.sodium, Sweat.potassium, Sweat.chloride,
    Sweat.sodium.left.arm, Sweat.sodium.right.arm, Sweat.potassium.left.arm,
    Sweat.potassium.right.arm, Sweat.chloride.left.arm,
    Sweat.chloride.right.arm, Diagnosis

@@

    Name = Patient.Name, Genotypes1 = Genotype.1, DOB = Birth.Date,
    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,

@@

    Name, DOB, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl, `Na:Cl`

@@

df$Diagnosis <- factor(df$Diagnosis,

@@

  labels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "CFTR-RD or CRMS", "Non-CF")

@@

df <- df[-c(
  78, 79, 183, 257, 259, 270, 314, 342, 691, 719, 724, 737, 812, 903, 1012,
  1066, 1126, 1151, 1202, 1209, 1222, 1224, 1340, 1428, 1456
), ]

@@

df$geno_risk <- factor(df$geno_risk,

@@

    "Mild/Mild", "Mild/Severe", "Mild/Unknown", "Severe/Mild", "Severe/Severe",
    "Severe/Unknown", "Unknown/Severe", "Unknown/Unknown", "Not CF/Not CF"

@@

df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")

@@

df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")

@@

df$geno_class <- factor(df$geno_class,

@@

    "F508del/F508del", "F508del/MF", "F508del/RF", "F508del/Unknown",
    "MF/F508del", "MF/MF", "MF/RF", "MF/Unknown", "RF/F508del", "RF/MF",
    "RF/Unknown", "Unknown/F508del", "Unknown/MF", "Unknown/Unknown",

@@

    "MF/MF", "MF/MF", "MF/RF", "MF/U", "MF/MF", "MF/MF", "MF/RF", "MF/U",
    "MF/RF", "MF/RF", "RF/U", "MF/U", "MF/U", "U", "Not CF"

@@

write.csv(df,

@@

  row.names = F, na = ""

@@

```{r results='asis'}
summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

@@

```{r results='asis'}
summary(tableby(`Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

@@

```{r results='asis'}
summary(tableby(`Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

@@

```{r}
# Fit a multiclass ROC
mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
roc_list <- list(
  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
)
ggroc(roc_list, legacy.axes = T) +
  theme_bw() +
  theme(legend.title = element_blank())
roc_table <- lapply(roc_list, function(r) {
  coords(r, "best",
    best.method = "youden",
    ret = c(
      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
      "tnr", "fnr"
    )
  )
})
roc_table <- do.call(rbind, roc_table)
kable(roc_table, digits = 3)
```

@@

```{r}
plot(roc_list[[2]],
  print.auc = T, print.thres = T,
  main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
)
plot(roc_list[[1]],
  print.auc = T, print.thres = T,
  main = "Cystic Fibrosis vs. Non-CF"
)
plot(roc_list[[3]],
  print.auc = T, print.thres = T,
  main = "CFTR-RD or CRMS vs. Non-CF"
)
```

>>>>

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
    "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
github_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Windows" = "C:/Users/Tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

chco <- read.csv(
    "./Data_Cleaned/chco_reidentified.csv",

@@

        "",
        "Test not performed",
        "Insufficient quant.",
        "repeat sweat chloride testing",
        "Please reorder testing and re",

@@

        PatientID,
        Name,
        DOB,
        Genotypes1,
        Genotypes2,
        Sex,
        Age.at.Test.in.Years,
        Na,
        K,
        Cl,
        Na_Left,
        Na_Right,
        K_Left,
        K_Right,
        Cl_Left,
        Cl_Right

@@

njh <- read.csv(
    "./Data_Cleaned/njh_full_list.csv",

@@

        "",
        "Test not performed",
        "Insufficient quant.",

@@

        Patient.ID,
        Patient.Name,
        Birth.Date,
        Genotype.1,
        Genotype.2,
        Sex,
        Age.at.test..in.years.,
        Sweat.sodium,
        Sweat.potassium,
        Sweat.chloride,
        Sweat.sodium.left.arm,
        Sweat.sodium.right.arm,
        Sweat.potassium.left.arm,
        Sweat.potassium.right.arm,
        Sweat.chloride.left.arm,
        Sweat.chloride.right.arm,
        Diagnosis

@@

        Name = Patient.Name,
        Genotypes1 = Genotype.1,
        DOB = Birth.Date,
        Genotypes2 = Genotype.2,
        Age = Age.at.test..in.years.,
        Na = Sweat.sodium,
        K = Sweat.potassium,
        Cl = Sweat.chloride,
        Na_Left = Sweat.sodium.left.arm,
        Na_Right = Sweat.sodium.right.arm,
        K_Left = Sweat.potassium.left.arm,
        K_Right = Sweat.potassium.right.arm,
        Cl_Left = Sweat.chloride.left.arm,

@@

        Name,
        DOB,
        Diagnosis,
        Genotypes1,
        Genotypes2,
        Sex,
        Age,
        Na,
        K,
        Cl,
        `Na:Cl`

@@

df$Diagnosis <- factor(
    df$Diagnosis,

@@

        "Cystic Fibrosis",
        "CFTR-RD or CRMS",
        "CFTR-RD or CRMS",
        "Non-CF"

@@

df <- df[
    -c(
        78,
        79,
        183,
        257,
        259,
        270,
        314,
        342,
        691,
        719,
        724,
        737,
        812,
        903,
        1012,
        1066,
        1126,
        1151,
        1202,
        1209,
        1222,
        1224,
        1340,
        1428,
        1456
]

@@

df$geno_risk <- factor(
    df$geno_risk,

@@

        "Mild/Mild",
        "Mild/Severe",
        "Mild/Unknown",
        "Severe/Mild",
        "Severe/Severe",
        "Severe/Unknown",
        "Unknown/Severe",
        "Unknown/Unknown",

@@

df$geno1_class <- replace(
    df$geno1_class,
    df$geno1_class %in% c("I", "II", "III"),
    "MF"
df$geno1_class <- replace(
    df$geno1_class,
    df$geno1_class %in% c("IV", "V"),
    "RF"

@@

df$geno2_class <- replace(
    df$geno2_class,
    df$geno2_class %in% c("I", "II", "III"),
    "MF"
df$geno2_class <- replace(
    df$geno2_class,
    df$geno2_class %in% c("IV", "V"),
    "RF"

@@

df$geno_class <- factor(
    df$geno_class,

@@

        "F508del/F508del",
        "F508del/MF",
        "F508del/RF",
        "F508del/Unknown",
        "MF/F508del",
        "MF/MF",
        "MF/RF",
        "MF/Unknown",
        "RF/F508del",
        "RF/MF",
        "RF/Unknown",
        "Unknown/F508del",
        "Unknown/MF",
        "Unknown/Unknown",

@@

        "MF/MF",
        "MF/MF",
        "MF/RF",
        "MF/U",
        "MF/MF",
        "MF/MF",
        "MF/RF",
        "MF/U",
        "MF/RF",
        "MF/RF",
        "RF/U",
        "MF/U",
        "MF/U",
        "U",
        "Not CF"

@@

write.csv(
    df,

@@

    row.names = F,
    na = ""

@@

```{r results='asis'}
summary(tableby(
    Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
    data = df,
    control = mycontrols
))
```

@@

```{r results='asis'}
summary(tableby(
    `Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
    data = df,
    control = mycontrols
))
```

@@

```{r results='asis'}
summary(tableby(
    `Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
    data = df,
    control = mycontrols
))
```

@@

```{r}
# Fit a multiclass ROC
mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
roc_list <- list(
    "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
    "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
    "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
)
ggroc(roc_list, legacy.axes = T) +
    theme_bw() +
    theme(legend.title = element_blank())
roc_table <- lapply(roc_list, function(r) {
    coords(
        r,
        "best",
        best.method = "youden",
        ret = c(
            "threshold",
            "sensitivity",
            "specificity",
            "accuracy",
            "tpr",
            "fpr",
            "tnr",
            "fnr"
        )
    )
})
roc_table <- do.call(rbind, roc_table)
kable(roc_table, digits = 3)
```

@@

```{r}
plot(
    roc_list[[2]],
    print.auc = T,
    print.thres = T,
    main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
)
plot(
    roc_list[[1]],
    print.auc = T,
    print.thres = T,
    main = "Cystic Fibrosis vs. Non-CF"
)
plot(
    roc_list[[3]],
    print.auc = T,
    print.thres = T,
    main = "CFTR-RD or CRMS vs. Non-CF"
)
```
